---
title: "DDS Case Study"
author: "Sid Swarupananda and Sean McWhirter"
date: "October 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries used

```{r}
library(plyr)
library(dplyr)
library(magrittr)
library(rlang)
library(ggplot2)
library(ggthemes)
library(knitr)
library(kableExtra)
library(class)
library(caret)
library(e1071)
library(ggthemes)
library(tidyverse)
library(GGally)
library(doBy)
library(ggmap)
library(maps)
library(openintro)
library(stringr)
```

# 1.	How many breweries are present in each state?

```{r}
# Read in beer and breweries data set using read.csv
Beers <- read.csv("https://raw.githubusercontent.com/BivinSadler/MDS-6306-Doing-Data-Science-Fall-2019/master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", header = TRUE, sep = ",")
Breweries <- read.csv("https://raw.githubusercontent.com/BivinSadler/MDS-6306-Doing-Data-Science-Fall-2019/master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv", header = TRUE, sep = ",")

detach(package:plyr)    
library(dplyr)

BrewPerState <-table(Breweries$State)
BrewPerState

Breweries %>% group_by(State) %>% summarize(count = n()) %>% ggplot(aes(x=reorder(State, count), y = count)) +
geom_bar(stat="identity", width=.5, fill="tomato3") +
labs(title="Ordered Bar Chart",
     x = "State",
     y = "Brewereis Count",
     subtitle="Breweries Per State Count",
     caption="source: Breweries.csv") +
     theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

# In the United States, range of breweries per state go from 1 brewery to 47 breweries. The average number of craft breweries per state is 10.9.

# Top States with the most breweries include:
# Colorado (47), California (39), Michigan (32), Oregon (29), Texas (28)
# Pennsylvania (25), Massachusetts (23), Washington (23), Indiana (22), and Wisconsin (20) are the next five in order
# Note: The top five states account for 31% of all breweries within the United States

# Bottom States
# North Dakota, South Dakota, West Virginia and District of Columbia are at the bottom with one brewery each.

# On another note, North Dakota, South Dakota and District of Columbia have a small population. But West Virginia is 38 in total population list in United States. If you are looking for an opportunity to expand into Beer market, West Virginia may be a good place to look at.

# 2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
# Renamed Brewery_id to Brew_ID to satisfy merging requirement
Beers <- rename(Beers, Brew_ID = Brewery_id)
BrewPub <- merge(Beers, Breweries, by="Brew_ID")

# Rename Name.x as Beer_Name and Name.y as Brewery_Name
BrewPub <- rename(BrewPub, Beer_Name = Name.x)
BrewPub <- rename(BrewPub, Brewery_Name = Name.y)

# First 6 observations of the merged file
kable(head(BrewPub, 6))

# Last 6 observations of the merged file
kable(tail(BrewPub,6))
```

# 3.   Address the missing values in each column.
```{r}
summary(BrewPub)
MissingValues <- sapply(BrewPub, function(x)sum(is.na(x)))
# Used kable library to make document more presentable
MissingValues %>% kable("html") %>% kable_styling()
```

# ABV and IBU are the only columns with missing values. However, since we have a high number of observations, we should be able to omit the missing values and still be able to determine any significant insights.  

# 4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
# Compute the median alcohol content unit for each state.
Alcohol <- BrewPub %>% na.omit() %>% group_by(State) %>% summarize(Median = median(ABV)) %>% arrange(Median)

# Compute the median international bitterness unit for each state.
Bitter <- BrewPub %>% na.omit() %>% group_by(State) %>% summarize(Median = median(IBU)) %>% arrange(Median)

# Plot a bar chart to compare ABV by state
ggplot(data=Alcohol, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="blue")+
  ggtitle("Median ABV by State") +
  labs(x="State",y="ABV")

# Plot a bar chart to compare IBU by state
ggplot(data=Bitter, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="blue")+
  ggtitle("Median IBU by State") +
  labs(x="State",y="IBU")
```


# Here we have 2 charts, which shows Medial alcohol content per state and Median international bitterness per state
# As you can see in both the charts the medial alcohol content per state is fairly consistent around 6% and median international bitterness per state is very much varied.

# 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
kable(BrewPub[which.max(BrewPub$ABV),])
kable(BrewPub[which.max(BrewPub$IBU),])
```

# We used which.max on our BrewPub dataset to get the maximum values in ABV and IBU columns 
# We identified Colorado as having the beer with the highest ABV, at 12.8% and we identified Oregon has having the beer with the highest IBU, at 138.

# 6.   Comment on the summary statistics and distribution of the ABV variable.
```{r}
BeerSummary <- (summary(BrewPub$ABV))
print(BeerSummary)
ggplot(BrewPub, aes(x=ABV)) + geom_histogram(color = "blue", fill = "white") + xlab("Alcohol by Volume")
```

# When we look at summary statistics of Alcohol by volume, we see that we have some beers don't even measure 1%. Median was 5.6%, Mean or average was 5.9% and the maximum ABV was around 13%. Also, note that we are missing 62 values in the dataset that we were provided

# Looking at a histogram of the data, it seems to be a bit right-skewed.  This makes sense because beer is not meant to be a very high ABV drink-the high ABV beers would be more rare.

# 7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}
# Draw a scatter plot to compare relationship between beer bitterness and alcohol content
ggplot(BrewPub, aes(x=IBU, y= ABV)) +
  geom_point(shape=1) +
   geom_smooth(method=lm) + # add linear regression line
ggtitle("Correlation between IBU and ABV") +
  labs(x="IBU - Bitterness of the beer",y="ABV - Alcoholic content of the beer")
# Pearson's product-moment correlation
cor.test(BrewPub$ABV,BrewPub$IBU)
```

# In this analysis, we examined data from two datasets regarding beers and breweries throughout the US. We examined the structure of this data, cleaned them, and combined them. We then performed various analyses, such as calculating median values for IBU and ABV by state, and determining which states had the beers with the highest ABV and IBU values. We finished by looking for a potential relationship between IBU and ABV, and found that there is evidence to suggest that a positive correlation exists between the two.

# The Pearson correlation coefficient, r, can take a range of values from +1 to -1. A value of 0 indicates that there is no association between the two variables. A value greater than 0 indicates a positive association; that is, as the value of one variable increases, so does the value of the other variable. Here we got the cor as 0.67 between ABV and IBU which suggests there  is a positive correlation.

# Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with "Ale" in its name other than IPA).

# Inida Pale Ale v. Other Type Ale
```{r}
BrewPub$BeerType <- ifelse(grepl("Ale$", BrewPub$Style, ignore.case = T), "Other Type Ale",
ifelse(grepl("IPA$", BrewPub$Style, ignore.case = T ), "India Pale Ale", "Other Type"))

BrewPub$BeerType = as.factor(BrewPub$BeerType)

BrewPub %>% ggplot(mapping=aes(x=Ounces, y=ABV, fill=BeerType)) + geom_boxplot() + 
  ggtitle("ABV by Ounces") + theme_economist()

# When looking at the beers by ounces, it appears that IPAs have higher median ABV values than other Ales. 
BrewPubIPAOther = BrewPub %>% filter(!is.na(BeerType) & !is.na(IBU) &!is.na(ABV)) %>%  
  filter(BeerType == "India Pale Ale" | BeerType == "Other Type Ale")
summary(BrewPubIPAOther)
BrewPubIPAOther = droplevels(BrewPubIPAOther, exclude = "Other Type")
summary(BrewPubIPAOther)

BrewPubIPAOther %>% ggplot(aes(x=IBU, ABV, color = BeerType)) + geom_point()

set.seed(6)
splitPerc = .75
trainIndices = sample(1:dim(BrewPubIPAOther)[1],round(splitPerc * dim(BrewPubIPAOther)[1]))
train = BrewPubIPAOther[trainIndices,]
test = BrewPubIPAOther[-trainIndices,]

BrewPubIPAOther %>% ggplot(aes(x = IBU, ABV, color = BeerType)) + geom_point() + geom_smooth() +
  ggtitle("ABV vs IBU") + theme_economist()
```

# Knn method to investigate the difference with respect to IBU and ABV between IPAs and Other Type of Ale

```{r}
# k = 3
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 3)
table(classifications,test$BeerType)
confusionMatrix(table(classifications,test$BeerType))

# k = 5
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 5)
table(test$BeerType,classifications)
confusionMatrix(table(test$BeerType,classifications))

# k = 10
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 10)
table(test$BeerType,classifications)
confusionMatrix(table(test$BeerType,classifications))


# k = 20
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 20)
table(test$BeerType,classifications)
CM = confusionMatrix(table(test$BeerType,classifications))
CM$overall[1]

## Loop for many k and one training / test partition
accs = data.frame(accuracy = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = i)
  table(test$BeerType,classifications)
  CM = confusionMatrix(table(test$BeerType,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}

plot(accs$k,accs$accuracy, type = "l", xlab = "k")
```

# Naive Bayes method to investigate the difference with respect to IBU and ABV between IPAs and Other Type of Ale

```{r}
model = naiveBayes(BrewPubIPAOther[,c(3,4)],BrewPubIPAOther$BeerType)
table(predict(model,BrewPubIPAOther[,c(3,4)]),BrewPubIPAOther$BeerType)
CM = confusionMatrix(table(predict(model,BrewPubIPAOther[,c(3,4)]),BrewPubIPAOther$BeerType))
CM

iterations = 100
masterAcc = matrix(nrow = iterations)
splitPerc = .7 #Training / Test split Percentage
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(BrewPubIPAOther)[1],round(splitPerc * dim(BrewPubIPAOther)[1]))
  train = BrewPubIPAOther[trainIndices,]
  test = BrewPubIPAOther[-trainIndices,]
  
  model = naiveBayes(train[,c(3,4)],as.factor(train$BeerType))
  table(predict(model,test[,c(3,4)]),as.factor(test$BeerType))
  CM = confusionMatrix(table(predict(model,test[,c(3,4)]),as.factor(test$BeerType)))
  masterAcc[j] = CM$overall[1]
}
MeanAcc = colMeans(masterAcc)
MeanAcc
```


#Other Insights
```{r}
library(usmap)
library(ggplot2)
library(mapproj)

BrewPubIPAOther$Ounces<-as.factor(BrewPubIPAOther$Ounces)

BrewPubIPAOther %>% ggplot(mapping=aes(x=Ounces, y=ABV, fill=BeerType)) + geom_boxplot() + 
  ggtitle("ABV by Ounces") + theme_economist()
```

```{r}
#Get Map Data
states <- map_data("state")

#Test plot of US
states %>% ggplot(mapping=aes(x=long, y=lat, group=group, fill=region)) + 
  geom_polygon(fill="white", color="black")

beerbybrew1 = BrewPub %>% filter(!is.na(ABV))

###Median of ABV by state###
abvmed = summaryBy(ABV~State, data=beerbybrew1, FUN=median)

#Get State names instead of abbreviations
abvmed1=abvmed
abvmed1$State<-as.character(abvmed1$State)

#Get rid of white space on left of abbreviation
abvmed1=abvmed1%>%mutate(State=str_trim(State))

#Do mutation on ABV percentage and getting state full name
abvmed2=abvmed1%>%mutate(State=abbr2state(State), ABV.median=ABV.median*100)
head(abvmed2)
#changing state to lowercase
abvmed3=abvmed2%>%mutate(State=tolower(State))
head(abvmed3)

#left_join to get long and lat for each stae
beerloc=left_join(states,abvmed3, by=c("region"="State"))

#Change ABV.Median to factor
beerloc$ABV.median<-as.numeric(beerloc$ABV.median)

#Plot ABV median by state
beerloc %>% ggplot(mapping=aes(x=long, y=lat, group=group, fill=ABV.median)) + 
  geom_polygon(color="black") + 
  coord_map(projection="albers", lat0=39, lat1=45) + 
  labs(title="Median ABV by State") + 
  theme_map() +
  labs(fill="Median ABV %")

```


#Heat Map of Breweries by State
```{r}
#Heat map for breweries by state
#Count the number of breweries per state
library(plyr)
brewst=count(Breweries$State)

#Get State names instead of abbreviations
brewst$x<-as.character(brewst$x)

#Get rid of white space on left of abbreviation
brewst1=brewst%>%mutate(x=str_trim(x))

#Do mutation on ABV percentage and getting state full name
brewst1=brewst1%>%mutate(state1=abbr2state(x))
head(brewst1)

#changing state to lowercase
brewst1=brewst1%>%mutate(state2=tolower(state1))
head(brewst1)
brewst2=brewst1%>%select(state2, freq)
head(brewst2)

#left_join to get long and lat for each stae
brewstate=left_join(states,brewst2, by=c("region"="state2"))
head(brewstate)

#Change ABV.Median to factor
brewstate$freq<-as.numeric(brewstate$freq)

#Plot ABV median by state
brewstate %>% ggplot(mapping=aes(x=long, y=lat, group=group,  fill=freq)) +
    geom_polygon(color="black") + 
    coord_map(projection="albers", lat0=39, lat1=45) + 
      labs(title="Number of Breweries by State") + 
      theme_map() + 
      labs(fill="# of Breweries")
```

