---
title: "DDS Case Study"
author: "Sid Swarupananda"
date: "October 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries used

```{r}
library(plyr)
library(dplyr)
library(magrittr)
library(rlang)
library(ggplot2)
library(ggthemes)
library(knitr)
library(kableExtra)
library(class)
library(caret)
library(e1071)
library(ggthemes)
```
# 1.	How many breweries are present in each state?

```{r}
# Read in beer and breweries data set using read.csv
Beers <- read.csv("https://raw.githubusercontent.com/BivinSadler/MDS-6306-Doing-Data-Science-Fall-2019/master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", header = TRUE, sep = ",")
Breweries <- read.csv("https://raw.githubusercontent.com/BivinSadler/MDS-6306-Doing-Data-Science-Fall-2019/master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv", header = TRUE, sep = ",")

BrewPerState <-table(Breweries$State)
BrewPerState

Breweries %>% group_by(State) %>% summarize(count = n()) %>% ggplot(aes(x=reorder(State, count), y = count)) +
geom_bar(stat="identity", width=.5, fill="tomato3") +
labs(title="Ordered Bar Chart",
     x = "State",
     y = "Brewereis Count",
     subtitle="Breweries Per State Count",
     caption="source: Breweries.csv") +
     theme(axis.text.x = element_text(angle=65, vjust=0.6))

```

# 2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
detach(package:plyr)    
library(dplyr)
# Renamed Brewery_id to Brew_ID to satisfy merging requirement
Beers <- rename(Beers, Brew_ID = Brewery_id)
BrewPub <- merge(Beers, Breweries, by="Brew_ID")

# Change variable names to more meaningful title
BrewPub <- full_join(Beers, Breweries, by="Brew_ID")

#kable - Used to create tables in LaTeX, HTML, Markdown and reStructuredText
# First 6 observations of the merged file
kable(head(BrewPub,6))

# Last 6 observations of the merged file
kable(tail(BrewPub,6))
```

# 3.   Address the missing values in each column.
```{r}
summary(BrewPub)
MissingValues <- sapply(BrewPub, function(x)sum(is.na(x)))
MissingValues %>% kable("html") %>% kable_styling()
BrewPub <- BrewPub %>% na.omit()
summary(BrewPub)
MissingValues <- sapply(BrewPub, function(x)sum(is.na(x)))
MissingValues %>% kable("html") %>% kable_styling()
```

# 4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
# Compute the median alcohol content unit for each state.
Firewater <- BrewPub %>% na.omit() %>% group_by(State) %>% summarize(Median = median(ABV)) %>% arrange(Median)

# Compute the median international bitterness unit for each state.
Bitter <- BrewPub %>% na.omit() %>% group_by(State) %>% summarize(Median = median(IBU)) %>% arrange(Median)

# Plot a bar chart to compare ABV by state
ggplot(data=Firewater, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_economist() + 
  scale_color_economist()+
  theme(axis.text.x=element_text(size=rel(0.8), angle=90)) +
  ggtitle("Median ABV by State") +
  labs(x="State",y="ABV")

# Plot a bar chart to compare IBU by state
ggplot(data=Bitter, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_economist() + 
  scale_color_economist()+
  theme(axis.text.x=element_text(size=rel(0.8), angle=90))+
  ggtitle("Median IBU by State") +
  labs(x="State",y="IBU")
```

# 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
kable(BrewPub[which.max(BrewPub$ABV),])
kable(BrewPub[which.max(BrewPub$IBU),])
```

# 6.   Comment on the summary statistics and distribution of the ABV variable.
```{r}
BeerSummary <- (summary(BrewPub$ABV))
print(BeerSummary)
```

# 7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}
# Draw a scatter plot to compare relationship between beer 
# bitterness and alcohol content
ggplot(BrewPub, aes(x=IBU, y= ABV)) +
  geom_point(shape=1) +
   geom_smooth(method=lm) + # add linear regression line
  theme_economist() + 
  scale_color_economist()+
  theme(axis.text.x=element_text(size=rel(1.0)))+
ggtitle("Correlation between IBU and ABV ") +
  labs(x="IBU",y="ABV")
# Pearson's product-moment correlation
cor.test(BrewPub$ABV,BrewPub$IBU)
```

# In this analysis, we examined data from two datasets regarding beers and breweries throughout the US. We examined the structure of this data, cleaned them, and combined them. We then performed various analyses, such as calculating median values for IBU and ABV by state, and determining which states had the beers with the highest ABV and IBU values. We finished by looking for a potential relationship between IBU and ABV, and found that there is evidence to suggest that a positive correlation exists between the two.

## Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with "Ale" in its name other than IPA).

```{r}
BrewPub$BeerType <- ifelse(grepl("Ale$", BrewPub$Style, ignore.case = T), "Other Type Ale",
ifelse(grepl("IPA$", BrewPub$Style, ignore.case = T ), "India Pale Ale", "Other Type"))

BrewPub$BeerType = as.factor(BrewPub$BeerType)

# Inida Pale Ale v. Other Type Ale
set.seed(6)
splitPerc = .75
BrewPubIPAOther = BrewPub %>% filter(BeerType == "India Pale Ale" | BeerType == "Other Type Ale")
summary(BrewPubIPAOther)
BrewPubIPAOther = droplevels(BrewPubIPAOther, exclude = "Other Type")
summary(BrewPubIPAOther)

BrewPubIPAOther <- BrewPubIPAOther %>% na.omit()

trainIndices = sample(1:dim(BrewPubIPAOther)[1],round(splitPerc * dim(BrewPubIPAOther)[1]))
train = BrewPubIPAOther[trainIndices,]
test = BrewPubIPAOther[-trainIndices,]

BrewPubIPAOther %>% ggplot(aes(x = IBU, ABV, color = BeerType)) + geom_point()
```

# Knn method to investigate the difference with respect to IBU and ABV between IPAs and Other Type of Ale

```{r}
# k = 3
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 3)
table(classifications,test$BeerType)
confusionMatrix(table(classifications,test$BeerType))

# k = 5
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 5)
table(test$BeerType,classifications)
confusionMatrix(table(test$BeerType,classifications))

# k = 10
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 10)
table(test$BeerType,classifications)
confusionMatrix(table(test$BeerType,classifications))


# k = 20
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = 20)
table(test$BeerType,classifications)
CM = confusionMatrix(table(test$BeerType,classifications))
CM$overall[1]

## Loop for many k and one training / test partition
accs = data.frame(accuracy = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train[,c(3,4)],test[,c(3,4)],train$BeerType, prob = TRUE, k = i)
  table(test$BeerType,classifications)
  CM = confusionMatrix(table(test$BeerType,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}

plot(accs$k,accs$accuracy, type = "l", xlab = "k")
```

# Naive Bayes method to investigate the difference with respect to IBU and ABV between IPAs and Other Type of Ale

```{r}
model = naiveBayes(BrewPubIPAOther[,c(3,4)],BrewPubIPAOther$BeerType)
table(predict(model,BrewPubIPAOther[,c(3,4)]),BrewPubIPAOther$BeerType)
CM = confusionMatrix(table(predict(model,BrewPubIPAOther[,c(3,4)]),BrewPubIPAOther$BeerType))
CM

iterations = 100
masterAcc = matrix(nrow = iterations)
splitPerc = .7 #Training / Test split Percentage
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(BrewPubIPAOther)[1],round(splitPerc * dim(BrewPubIPAOther)[1]))
  train = BrewPubIPAOther[trainIndices,]
  test = BrewPubIPAOther[-trainIndices,]
  
  model = naiveBayes(train[,c(3,4)],as.factor(train$BeerType))
  table(predict(model,test[,c(3,4)]),as.factor(test$BeerType))
  CM = confusionMatrix(table(predict(model,test[,c(3,4)]),as.factor(test$BeerType)))
  masterAcc[j] = CM$overall[1]
}
MeanAcc = colMeans(masterAcc)
MeanAcc
```

